A Comparsion of Traffic Engineering Methods for Application Perceived Metrics

### 2015-07-14 ncdnpaper/Evaluation2-alt.tex
first complete version after fixing most of the stuff

The experiments so far  suggest that a \planned\ scheme that engineers placement and routing once a day based on the previous day's demand performs poorly compared to an \unplanned\ scheme, \invlru. Therefore, in this section, we evaluate the performance of two alternative \planned\ schemes.

@@

Executing a \planned\ placement requires considerable effort---measuring content matrix, solving a computationally intensive optimization, and moving content to new locations. Further, a \planned\ placement needs to be executed  8 times a day (or possibly more) even to match the cost achieved by an \unplanned\ strategy. Our position is that NCDNs are better served by opting for a much simpler \unplanned\ strategy and provisioning more storage, in which case, an \unplanned\ strategy  already obtains  a network cost close to the best a \planned\ strategy can possibly achieve.

@@

The latency cost of \invlru\ relative to \optrpfuturel\ improves with an increase in storage ratio. At the smallest storage ratio, \invlru\  has 70-110\% higher latency cost than \optrpfuturel. The difference reduces to  14-34\% at the maximum storage ratio. Higher storage ratio translate to higher cache hit rates, which reduces propagation delay of transfers and lowers link utilizations. Both these factors contribute to  a smaller latency cost for \invlru. 
This finding shows that NCDNs can achieve close to best latency costs with an \unplanned\ scheme \invlru\ and provisioning moderate amounts of storage.

@@

\subsection{Effect of NCDN traffic on network cost}
This experiment, unlike previous experiments, considers a network consisting of both ISP and NCDN traffic. Our goal is to evaluate how network costs change as the fraction of NCDN traffic increases in the network. Second, we seek to examine the benefit of optimizing routing over an unplanned routing scheme, InverseCap. To this end, we compare the performance of \invlru\ and \optlru\ schemes. The latter scheme optimizes routing for the combined traffic matrix due to NCDN traffic and ISP transit traffic. 
In order to estimate the best gains achievable with an optimized routing, we provide to the \optlru\ scheme knowledge of future ISP traffic matrices. \optlru\ cannot be provided the knowledge of future NCDN traffic matrices because NCDN traffic matrices can only be measured from experiment itself and we do not know them beforehand. We optimize routing once a day in this experiment. Varying the frequency of routing update did not improve \optlru's performance.

@@

We present the results of our experiments on the news trace in Figure \ref{fig:ispncdn}. We vary the fraction of NCDN to ISP traffic, and report MLUs normalized by the total volume of ISP and NCDN traffic.
Our results are not independent of the scale of simulations: a larger or a smaller scaling of CDN trace may give quantitatively different conclusions. Hence, we only make qualitative conclusions from this experiment. First, we find that as the fraction of NCDN traffic increases, MLU decreases for both schemes. This is intuitive since a large fraction of NCDN traffic is served from caches located at PoPs. Second, as NCDN traffic increases optimizing routing (\optlru) gives lesser benefits compared to InverseCap routing.  In a network dominated by NCDN traffic, optimizing routing gives almost no benefits over \invlru. We find these results to be consistent with our earlier experiments with NCDN traffic only.

>>>>

The experiments so far  suggest that a \planned\ scheme that engineers placement and routing once a day based on the previous day's demand performs poorly compared to a \unplanned\ scheme, \invlru. Therefore, in this section, we evaluate the performance of two alternative \planned\ schemes.

@@

Executing a \planned\ placement requires considerable effort---measuring content matrix, solving a computationally intensive optimization, and moving content to new locations. Further, a \planned\ placement needs to be executed  8 times a day (or possibly more) even to match the cost achieved by a \unplanned\ strategy. Our position is that NCDNs are better served by opting for a much simpler \unplanned\ strategy and provisioning more storage, in which case, a \unplanned\ strategy  already obtains a network cost close to the best a \planned\ strategy can possibly achieve.

@@

The latency cost of \invlru\ relative to \optrpfuturel\ improves with an increase in storage ratio. At the smallest storage ratio, \invlru\  has 70-110\% higher latency cost than \optrpfuturel. The difference reduces to  14-34\% at the maximum storage ratio. Higher storage ratio translate to higher cache hit rates, which reduces propagation delay of transfers and lowers link utilizations. Both these factors contribute to  a smaller latency cost for \invlru. 
This finding shows that NCDNs can achieve close to best latency costs with a \unplanned\ scheme \invlru\ and provisioning moderate amounts of storage.

@@

\subsection{Effect of NCDN traffic on network cost}
This experiment, unlike previous experiments, considers a network consisting of both ISP and NCDN traffic. Our goal is to evaluate how network costs change as the fraction of NCDN traffic increases in the network. Second, we seek to examine the benefit of optimizing routing over an unplanned routing scheme, InvCap. To this end, we compare the performance of \invlru\ and \optlru\ schemes. The latter scheme optimizes routing for the combined traffic matrix due to NCDN traffic and ISP transit traffic. 
In order to estimate the best gains achievable with an optimized routing, we provide to the \optlru\ scheme knowledge of future ISP traffic matrices. \optlru\ cannot be provided the knowledge of future NCDN traffic matrices because NCDN traffic matrices can only be measured from experiment itself and we do not know them beforehand. We optimize routing once a day in this experiment. Varying the frequency of routing update did not improve \optlru's performance.

@@

We present the results of our experiments on the news trace in Figure \ref{fig:ispncdn}. We vary the fraction of NCDN to ISP traffic, and report MLUs normalized by the total volume of ISP and NCDN traffic.
Our results are not independent of the scale of simulations: a larger or a smaller scaling of CDN trace may give quantitatively different conclusions. Hence, we only make qualitative conclusions from this experiment. First, we find that as the fraction of NCDN traffic increases, MLU decreases for both schemes. This is intuitive since a large fraction of NCDN traffic is served from caches located at PoPs. Second, as NCDN traffic increases optimizing routing (\optlru) gives lesser benefits compared to InvCap routing.  In a network dominated by NCDN traffic, optimizing routing gives almost no benefits over \invlru. We find these results to be consistent with our earlier experiments with NCDN traffic only.

### 2015-07-14 ncdnpaper/LinearPrograms.tex
first complete version after fixing most of the stuff

\section{NCDN management strategies}
\label{sec:optimize}

@@

To answer the above questions, we develop an optimization model for NCDNs to jointly optimize placement, routing, and redirection so as to optimize network cost or user-perceived latency. We formulate the optimization problem as a mixed-integer program (MIP), present hardness and inapproximability results, and discuss approximation heuristics to solve MIPs for realistic problem sizes.

@@

\subsection{Computational hardness}
\optloc\ is the decision version of the NCDN problem. The proofs for these theorems are presented in a techreport  \cite{techreportNCDN}.
%We show that the NCDN-problem is NP-Complete and is inapproximable to within an exponential factor.

>>>>

\section{NCDN joint optimization}
\label{sec:optimize}

@@

We develop an optimization model for NCDNs to jointly optimize placement, routing, and redirection so as to optimize network cost or user-perceived latency. We formulate the optimization problem as a mixed-integer program (MIP), present hardness and inapproximability results, and discuss approximation heuristics to solve MIPs for realistic problem sizes.

@@

\subsection{Computational hardness}
\optloc\ is the decision version of the NCDN problem. The proofs for these theorems are presented in Appendix \ref{sec:npcreduction}.

%We show that the NCDN-problem is NP-Complete and is inapproximable to within an exponential factor.

### 2015-07-14 tepaper/section1.tex
intro and background revision: incorporated comments until background 2.1
first complete version after fixing most of the stuff

\chapter{Traffic Engineering: An Application-Centric Comparison}
\label{ch:beyondmlu}
Traditionally, traffic engineering (TE) has been studied as an optimization problem that takes as input a traffic matrix (TM) and seeks to compute routes so as to minimize a network cost function. The cost function is intended to capture the severity of congestion hotpsots based on link utilization levels. For example, the most widely used cost function, MLU, is simply the utilization of the most utilized link in the network \cite{COPE,TEXCP,MultiTM,Cohen}; others  sum over all links a convex function of their utilization  (so as to penalize highly utilized links more)  \cite{fortz2000internet,fortz2002traffic}. There are two implicit assumptions underlying this line of work. First, maintaining low link utilization improves user-perceived application performance under typical load conditions. Second, maintaining low link utilization increases the effective capacity of the network by enabling it to accommodate unexpected surges in the traffic demand.

@@

More surprisingly, we find that application adaptation to location diversity,  i.e., the ability to download content from multiple potential locations, blurs differences even in the achieved capacities of different TE schemes enabling all of them to be near-optimal. With location diversity, we find that the inverse of the MLU is no longer a meaningful metric of capacity. Instead, we formalize a new metric of the capacity achieved by a TE scheme called the {\em surge protection factor} (SPF) that captures the factor of increase in demand that can be sustained while accounting for location diversity. TE schemes calculate routing based on a measured traffic matrix to achieve a desired network cost, but application adaptation to location diversity changes the traffic matrix itself in response to a change in routing resulting a different network cost than expected. As a result, the optimal TE scheme with perfect knowledge of traffic matrix and sub-optimal TE schemes like OSPF weight-tuning \cite{fortz2000internet} end-up achieving the same SPF. Even the static routing scheme, \invcap, achieves an SPF at most 30\% worse than optimal TE.

The rest of the chapter is as follows. Section \ref{sec:locdiv-background} explains how location diversity changes the TE problem. Section \ref{sec:exp_setup} presents our simulation setup. Section \ref{sec:app_performance} compares the application performance of TE schemes and Section \ref{sec:capacity} compares their achieved capacity under location diversity.

>>>>

\chapter{An application-centric comparison of ISP traffic engineering schemes}
\label{ch:beyondmlu}

Traffic engineering (TE) techniques are used by Internet Service Providers (ISPs) for computing network routing based on expected traffic demands. Our work focus on two aspects of ISP traffic engineering that have received less attention in prior work and are of key importance in a content-dominated network. First, the impact of traffic engineering on user-perceived peformance such as TCP file download times. Second, the impact of application adaptation on traffic engineering. In particular we focus on a form of adaptation enabled by \emph{location diversity}, or the ability of applications to download content from multiple locations. Location diversity is prevalent in the Internet due to several applications and services such as content delivery networks, peer-to-peer applications and mirrored websites.

While ISP traffic engineering schemes are commonly evaluated using link-utilization based metrics, these metrics are ill-suited to compare schemes based on user-perceived performance or their network capacity, i.e., the factor of surges in traffic demand that a network can tolerate, under location diversity. As discussed previously in Chapter \ref{ch:te-background}, link utilization-based metrics may not reflect application performance that depends on other metrics such as propagtion delay, access link capacity and backbone link capacity also. Further, link utilization based metrics, in particular MLU, do not reflect network capacity under location diversity as explained before.

There are two key parts of our approach to evaluate a TE scheme. First, we carefully and at scale simulate network traffic as a collection of TCP flows. This empirical approach  enables us to accurately measure application performance metrics --- TCP throughput for elastic traffic and a quality-of-service metric,  MOS score for VoIP quality \cite{MOS-formula}, for inelastic traffic --- and model application adaptation to location diversity, e.g., how an application splits traffic among multiple content locations. Second, we use a novel capacity metric called Surge Protection Factor (SPF). Unlike MLU, SPF captures the factor of increase in demand that can be sustained while accounting for location diversity.

We show following results from comparing user-perceived performance of TE schemes based on real network topologies and traffic demands. All state-of-the-art TE schemes achieve nearly identical application performance at typical Internet load levels. In fact, even a demand-oblivious TE scheme --- static shortest-path routing with link weights inversely proportional to the capacity (\invcap) (i.e., no engineering at all) --- achieves the same application performance as optimal TE. Ironically, demand-aware TE schemes that engineer for unexpected traffic spikes (e.g., COPE \cite{COPE}) consistently hurt TCP throughput despite achieving near-optimal MLU. Overall, these results show that link utilization based metrics are poorly correlated with user-perceived performance.

We show following results from comparing SPF achieved by TE schemes. The optimal TE scheme with perfect knowledge of traffic matrix and sub-optimal TE schemes like demand-aware TE by optimizing OSPF link weights \cite{fortz2000internet} end-up achieving the same SPF. Even the demand-oblivious scheme, \invcap, achieves an SPF at most 30\% worse than optimal TE. Overall, these results show that application adaptation to location diversity blurs differences between traffic engineering schemes.

The rest of the chapter is as follows. Section \ref{sec:locdiv-background} introduces location diversity and proposes a new capacity metric SPF to compare traffic engineering schemes. Section \ref{sec:exp_setup} presents our simulation setup. Section \ref{sec:app_performance} compares the application performance of TE schemes and Section \ref{sec:capacity} compares their achieved capacity under location diversity.

%Our approach to evaluating traffic engineering is empirical and application-centric.  Our experiments carefully and at scale simulate end-to-end application behavior. This approach enables us to evaluate TE schemes with respect to their impact on application performance in terms of TCP throughput for elastic traffic and a quality-of-service metric,  MOS score for VoIP quality \cite{MOS-formula}, for inelastic traffic. Further, this approach enables us to model application adaptation to location diversity and measure capacity using a novel capacity metric called Surge Protection Factor (SPF). SPF captures the factor of increase in demand that can be sustained while accounting for location diversity.

%Further, this approach enables us to simulate application adaptation to location diversity.

%Our research shows that link utilization based metrics, in particular the maximum link utilization (MLU) is a poor metric to evaluate user-perceived performance or \emph{network capacity}, i.e., the maximum surge in traffic demand that a network can support. Our results support our previous hypothesis that link utilization based metrics may not reflect user-perceived performance that depends on other factors also.
%Previously, we explained why link utilization based metrics may not reflect user-perceived performance that depends on other factors also.  Previously, we explained why MLU is no longer an effective metric of capacity in the presence of location diversity. In this Chapter, we propose a new capacity metric called Surge Protection Factor (SPF) for comparing traffic engineering schemes. SPF captures the factor of increase in demand that can be sustained while accounting for location diversity.

\eat{
	Traditionally, traffic engineering (TE) has been studied as an optimization problem that takes as input a traffic matrix (TM) and seeks to compute routes so as to minimize a network cost function. The cost function is intended to capture the severity of congestion hotpsots based on link utilization levels. For example, the most widely used cost function, MLU, is simply the utilization of the most utilized link in the network \cite{COPE,TEXCP,MultiTM,Cohen}; others  sum over all links a convex function of their utilization  (so as to penalize highly utilized links more)  \cite{fortz2000internet,fortz2002traffic}. There are two implicit assumptions underlying this line of work. First, maintaining low link utilization improves user-perceived performance under typical load conditions. Second, maintaining low link utilization increases the effective capacity of the network by enabling it to accommodate unexpected surges in the traffic demand.

@@

	More surprisingly, we find that application adaptation to location diversity,  i.e., the ability to download content from multiple potential locations, blurs differences even in the achieved capacities of different TE schemes enabling all of them to be near-optimal. With location diversity, we find that the inverse of the MLU is no longer a meaningful metric of capacity. Instead, we formalize a new metric of the capacity achieved by a TE scheme called the {\em surge protection factor} (SPF) that captures the factor of increase in demand that can be sustained while accounting for location diversity. TE schemes calculate routing based on a measured traffic matrix to achieve a desired network cost, but application adaptation to location diversity changes the traffic matrix itself in response to a change in routing resulting a different network cost than expected. As a result, the optimal TE scheme with perfect knowledge of traffic matrix and sub-optimal TE schemes like OSPF weight-tuning \cite{fortz2000internet} end-up achieving the same SPF. Even the static routing scheme, \invcap, achieves an SPF at most 30\% worse than optimal TE.
}
